---
title: "Project Proposal"
author: "Seokjun Choi and Sho Kawano"
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation: 
    theme: serif
    incremental: false
    reveal_options:
      slideNumber: T
---

```{r, include=FALSE}
library(tidyverse)
# library(gridExtra)
data <- read.csv("MH_survey_only_higher_index.csv",
    na.strings = "NA")
names(data)
names(data)[1] <- "gender"

data <- data[, -c(17:22)]


# Assigning factors
data$gender <- factor(data$gender)
data$age_group <- factor(data$age_group)
data$country_lockdown <- factor(data$country_lockdown)
data$marital <- factor(data$marital)
data$smoking <- factor(data$smoking)
data$fivfruitveg <- factor(data$fivfruitveg)
data$shielded <- factor(data$shielded)
data$week_soc_distancing <- factor(data$week_soc_distancing)
data$athlete <- factor(data$athlete)
```



# Changed Dataset

* Similar but different research questions
* Examine how athletes & non-athletes were affected by COVID-19.
* Online Survey Data Collected as the UK + Ireland emerged from COVID-19 Lockdowns 
  + Taken 14-16 weeks after lockdowns. June/July 2020. 
  + n=753 (# of complete cases: 457)


# Data Description: Factors

* Gender (Binary). <small>Male/Female</small>
* Shielded (Quarantined). <small>Yes/No </small>
* Athlete/Non-Athlete
* Marital Status  <small>5-level </small>
* Age group <small>7-level & ordered. </small>
* Country during lockdown. <small>7-level.</small>
* Smoking status. <small>7-level.</small>
* Week Social Distancing. <small>7-level & ordered.</small>
* \# in Lockdown bubble. <small>7-level & ordered.</small>



# Data Description: Numeric

* Hours Sleep.  <small>Range: 1.5-10.5 hours</small>
* Athlete Identification Index (AIMS). <small>Range: 1-7.</small>
* Hospital Anxiety & Depression Scale (HADS). <small>Range: 0-36.</small>
* Brief Resilience Scale  (RES). <small>Range: 6-30</small>
* Loneliness Scale (LONE) <small>Range: 0-6</small>
* *Response:* MHC-SF. Represents overall well-being. <small>Range: 2-70</small>



# Main Research Questions 

* Were athletes and non-athletes affected from COVID-19 lockdown/social distancing in the same way?
* Are there sex-based differences in the effects of COVID-19 on mental health?
* Under controlled demographic information, what is the relationship between MH-SF index and other sub-indices in this survey?
* Are there additional variables having strong relationship with MH-SF index?


# EDA

MHC_SF VS QUARANTINED (WITH GENDER)
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
    mutate(gender = ifelse(gender==1, "Male", "Female"), 
           athlete = ifelse(athlete==1, "Athlete", "Non-Athlete")) %>% 
    mutate(Qurantined=shielded) %>% 
    ggplot(aes(x=athlete, y=MHC_SF_OVERALL, color=Qurantined)) + 
    geom_boxplot() + 
    facet_grid(~gender) + xlab("") + theme_bw() #doesn't error occur?
```

#

MHC_SF VS HADS (GENDER - ATHLETE)
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
    mutate(gender = ifelse(gender==1, "Male", "Female"), 
           athlete = ifelse(athlete==1, "Athlete", "Non-Athlete")) %>% 
    ggplot(aes(x=HADS_OVERALL, y=MHC_SF_OVERALL)) + geom_point(aes(color=athlete)) + 
    facet_grid(athlete~gender) + theme_bw() +
    geom_smooth(se=FALSE)
```

#

MHC_SF VS SOC_DIST (GENDER - ATHLETE)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
    mutate(gender = ifelse(gender==1, "Male", "Female"), 
           athlete = ifelse(athlete==1, "Athlete", "Non-Athlete")) %>% 
    ggplot(aes(x=week_soc_distancing, y=MHC_SF_OVERALL)) + geom_boxplot(aes(color=athlete)) + 
    facet_grid(athlete~gender) + theme_bw() +
    geom_smooth(se=FALSE)
```

#

MHC_SF VS AGE_GROUP (GENDER - ATHLETE)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
    mutate(gender = ifelse(gender==1, "Male", "Female"), 
           athlete = ifelse(athlete==1, "Athlete", "Non-Athlete")) %>% 
    ggplot(aes(x=age_group, y=MHC_SF_OVERALL)) + geom_boxplot(aes(color=athlete)) + 
    facet_grid(athlete~gender) + theme_bw() +
    geom_smooth(se=FALSE)
```

#

MHC_SF VS LONE_TOTAL (GENDER - ATHLETE)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
    mutate(gender = ifelse(gender==1, "Male", "Female"), 
           athlete = ifelse(athlete==1, "Athlete", "Non-Athlete")) %>% 
    ggplot(aes(x=LONE_TOTAL, y=MHC_SF_OVERALL)) + geom_point(aes(color=athlete)) + 
    facet_grid(athlete~gender) + theme_bw() +
    geom_smooth(se=FALSE)
```


# Collinearity 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(corrplot)
library(RColorBrewer)

cor_tbl = data %>% select(-MHC_SF_OVERALL) %>%  
    mutate_each(as.numeric)  %>% na.omit() %>% cor()

corrplot(cor_tbl, type="upper", order="hclust", method="color")
```

#

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% ggplot(aes(x=RES_TOTAL, y=HADS_OVERALL)) + geom_point() + geom_smooth(se=FALSE)+ theme_bw()
```

#

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% ggplot(aes(x=athlete, y=AIMS_TOTAL)) + geom_point(position="jitter") + 
  geom_boxplot(alpha=0.01, color="blue")+ theme_bw()
```


# model selection

- stepwise model-selection
    - using 'step' function in R

- Removing variables with Collinearity Issues
    - AIMS & athlete: 'AIMS' is deleted by the stepwise selection
    - RES & HADS: using AIC criterion, we delete 'RES'

- possible interaction?
    - Not exist. (through F-test)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
final_model = lm(formula = MHC_SF_OVERALL ~ gender + age_group + week_soc_distancing + 
    athlete + HADS_OVERALL + LONE_TOTAL, data = data[complete.cases(data), ])
model_interaction = lm(formula = MHC_SF_OVERALL ~ (gender + age_group + week_soc_distancing + 
    athlete + HADS_OVERALL + + LONE_TOTAL)^2, data = data[complete.cases(data), ])
anova(final_model, model_interaction)
```



# Model

<small> 
\begin{align*}
y_{i} = &\mu 
+ \beta_{1}I_{gender=2}(x_i)
+ \sum_{j=2}^7\beta_{2j}I_{\text{age grp}=j}(x_{2i}) \\
&+ \sum_{j=1}^7 \beta_{9_j}I_{\text{wk soc dist}=j}(x_{9i})
+ \beta_10 I_{\text{athlete}=2}(x_{10i}) \\
&+ \beta_{13} x_{13i}
+ \beta_{15} x_{15i}
+ \epsilon_i \\
\end{align*}

where $\epsilon ~ N(0,\sigma^2)$, and

- $I_A$s are indicator variables ($I_A(x)= 1$ if $x \in A$, or $0$ otherwise)
- $x_{13i}$: HADS of i-th point
- $x_{15i}$: LONE

</small>




# fitting

```{r, echo=FALSE, warning=FALSE, message=FALSE}
final_model = lm(formula = MHC_SF_OVERALL ~ gender + age_group + week_soc_distancing + 
    athlete + HADS_OVERALL + LONE_TOTAL, data = data[complete.cases(data), ])
final_model %>% summary()
```

#

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(car)
```
```{r}
final_model %>% Anova()
```


# diagnosis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow = c(2, 2))
plot(final_model)
```

#

For the high residual point,

```{r, echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow = c(3, 2))
for(i in (names(final_model$model)[-1])) {
    now_data <- as.numeric(data[, i])
    if (i == "week_soc_distancing") {
      now_data <- now_data - 1
    }
    boxplot(now_data, xlab = i, horizontal = TRUE)
    points(now_data[364], 1, pch = 19, col = "red", cex = 2)
}
```

#

To check the heteroskedasticity,

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data_without_na <- data[complete.cases(data), ]
par(mfrow = c(1, 1))
n_lab_to_boxplot <- function(factor_var, y=1) {
    lev <- levels(factor_var)
    for(i in seq_along(lev)) {
        text(i, y, paste("n=", sum(factor_var == lev[i]), sep = ""))
    }
}
plot(final_model$residuals ~ data_without_na$week_soc_distancing)  # <- hmm. but boxes are similar
n_lab_to_boxplot(data_without_na$week_soc_distancing, y = 25) #we can guess what makes the problem

```


#

(leverage)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow = c(1, 1)); plot(final_model, which = 4)
```



# Grouped Lasso model
<small>
\begin{align*}
y_{i} = &\mu 
+ \beta_{1}I_{gender=2}(x_i)
+ \sum_{j=2}^7\beta_{2j}I_{\text{age grp}=j}(x_{2i})
+ \sum_{j=2}^7\beta_{3j}I_{\text{country lockdown}=j}(x_{3i}) \\
&+ \sum_{j=2}^5\beta_{4j}I_{\text{marital}=j}(x_{4i})
+ \sum_{j=2}^7\beta_{5j}I_{\text{smoking}=j}(x_{5i})
+ \beta_{6}I_{\text{fruits}=2}(x_{6i}) \\
&+ \beta_{7} x_{7i}
+ \beta_{8}I_{\text{shielded}=2}(x_{8i}) 
+ \sum_{j=1}^7 \beta_{9_j}I_{\text{wk soc dist}=j}(x_{9i})
+ \beta_{10} I_{\text{athlete}=2}(x_{10i}) \\
&+ \beta_{11} x_{11i}
+ \beta_{12} x_{12i}
+ \beta_{13} x_{13i}
+ \beta_{14} x_{14i}
+ \beta_{15} x_{15i}
+ \epsilon_i
\\
\end{align*}

where 

- $I_A$s are indicator variables, ($I_A(x)= 1$ if $x \in A$, or $0$ otherwise)
- $x_{7i}$: hour_sleep of the i-th observation
- Likewise, $x_{11i}$: lockdown_bubble, $x_{12i}$: AIMS, $x_{13i}$: HADS, $x_{14i}$: RES, and $x_{15i}$: LONE

with an error term $\epsilon_i$.

</small>



# Grouped Lasso: fitting

- Setting each category and each continuous variable to its own group.
- Notation
    - \(\mathbf{X}^{(g)} = (\mathbf{X}_{g1}, \mathbf{X}_{g2}, ..., \mathbf{X}_{gp_g})\)
    - \(\beta^{(g)} = (\beta_{g1}, \beta_{g2}, ..., \beta_{gp_g})^T.\)

Then our objective function for grouped LASSO is:
\[ \big | \big|y-\sum_{g=1}^G \mathbf{X}^{(g)}\beta^{(g)} \big |\big|_2^2 + \lambda \sum_{g=1}^G \sqrt{p_g}||\beta^{(g)}||_{2}\]


# Choosing $\lambda$
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# group lasso from full model
library(gglasso)
# ?gglasso

model_full <- lm(MHC_SF_OVERALL ~ ., data = data)
design_matrix_full <- model.matrix(model_full)[, -1]
conti_var_idx <- c(24, 33, 35, 36, 37, 38)
design_matrix_full <- scale(model.matrix(model_full)[, -1])
# for(i in conti_var_idx) {
#     design_matrix_full[, i] <- scale(design_matrix_full[, i])
# }
response_full <- scale(model_full$model$MHC_SF_OVERALL)

model_full_col_group <- c(1, 2, 2, 2, 2, 2, 2,
    3, 3, 3, 3, 3, 4, 4, 4, 4, 5, 5, 5, 5, 5, 5, 6, 7, 8,
    9, 9, 9, 9, 9, 9, 9, 10, 11, 12, 13, 14, 15)
# data.frame(colnames(design_matrix_full), model_full_col_group) #for check
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# fit with cv
model_full_lasso_cv <- cv.gglasso(
    design_matrix_full, response_full,
    group = model_full_col_group,
    pred.loss = "L2",
    nfolds = 10)
```

With 10-fold CV,

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot(model_full_lasso_cv)
abline(v = log(0.03), col = "blue")
text(log(0.03), 1, "lambda: 0.03", pos = 4, col = "blue")
```



# Grouped Lasso: fitting result
```{r}
library(gglasso)
model_full_lasso_fit <- gglasso(
    design_matrix_full, response_full,
    lambda = 0.03,
    group = model_full_col_group, loss = "ls")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
coeff_lasso <- round(coef(model_full_lasso_fit), 3)
```

#
```{r}
coeff_lasso[1:20, ] %>% as.data.frame()
```

#
```{r}
coeff_lasso[21:39, ] %>% as.data.frame()
```

#

The grouped lasso gives the model with 20 nonzero coefficients:
<small>
\begin{align*}
y_{i} = &\mu 
+ \beta_{1}I_{gender=2}(x_i)
+ \sum_{j=2}^7\beta_{2j}I_{\text{age grp}=j}(x_{2i})
+ \beta_{6}I_{\text{fruits}=2}(x_{6i})
+ \sum_{j=1}^7 \beta_{9_j}I_{\text{wk soc dist}=j}(x_{9i}) \\
&+ \beta_{10} I_{\text{athlete}=2}(x_{10i})
+ \beta_{13} x_{13i}
+ \beta_{14} x_{14i}
+ \beta_{15} x_{15i}
+ \epsilon_i
\\
\end{align*}

where 

- $I_A$s are indicator variables, ($I_A(x)= 1$ if $x \in A$, or $0$ otherwise)
- $x_{13i}$: HADS
- $x_{14i}$: RES
- $x_{15i}$: LONE

with an error term $\epsilon_i$.


Or,
MHC_SF_OVERALL ~ gender + age_group +
    fivfruitveg + week_soc_distancing +
    athlete + HADS_OVERALL + RES_TOTAL + LONE_TOTAL


</small>

#
```{r, echo=FALSE, warning=FALSE, message=FALSE}

try_log_lambdas <- seq(-2.4, -4, by = -0.01)
coeff_mat <- matrix(0, 38, length(try_log_lambdas))
rownames(coeff_mat) <- colnames(design_matrix_full)
colnames(coeff_mat) <- try_log_lambdas
for (i in seq_len(length(try_log_lambdas))) {
    log_lambda <- try_log_lambdas[i]
    beta_at_each_lambda <- gglasso(
        design_matrix_full, response_full,
        lambda = exp(log_lambda),
        group = model_full_col_group, loss = "ls")$beta
    coeff_mat[, i] <- beta_at_each_lambda
}


#second plot
plot(try_log_lambdas, rep(0, length(try_log_lambdas)), type = "n",
    xlim = c(min(try_log_lambdas), max(try_log_lambdas)), ylim = c(-0.4, 0.2),
    xlab = "log-lambda", ylab = "coefficients")
for (i in 1:38) {
    lines(try_log_lambdas, coeff_mat[i, ], col = i + 1)
    coeff_at_eval_point <- coeff_mat[i, which(try_log_lambdas == -3.5)]
    if (coeff_at_eval_point > 0.02 | coeff_at_eval_point < -0.02) {
        text(-4, coeff_mat[i, length(try_log_lambdas)],
            rownames(coeff_mat)[i], pos = 4)
        # print(rownames(coeff_mat)[i])
    }
}
abline(h = 0)
abline(v = log(0.03), lty = 2)
text(-3.5, -0.2, "lambda=0.03")

```

# OLS vs Grouped Lasso

```{r, echo=FALSE, warning=FALSE, message=FALSE}
lasso_coeff <- coef(model_full_lasso_fit)[coef(model_full_lasso_fit) != 0]

after_lasso_lm_fit_non_scaled <- lm(MHC_SF_OVERALL ~ gender + age_group +
    fivfruitveg + week_soc_distancing +
    athlete + HADS_OVERALL + RES_TOTAL + LONE_TOTAL, data = data)
scaled <- scale(model.matrix(after_lasso_lm_fit_non_scaled)[, -1])
response_al <- scale(after_lasso_lm_fit_non_scaled$model$MHC_SF_OVERALL)
after_lasso_lm_fit <- lm(response_al ~ scaled)
# summary(after_lasso_lm_fit)

# coef(after_lasso_lm_fit)[coef(after_lasso_lm_fit) != 0]
round(data.frame(lm_coeff = coef(after_lasso_lm_fit),
    lasso_coeff = lasso_coeff), 3)


# par(mfrow = c(2, 2))
# plot(after_lasso_lm_fit)

```


# Discussion 

Any questions? Comments? Suggestions? 

